Think deep for reef resilience

### 2025-08-16 paper/paper_draft.qmd
Tidy up bioregion scale

## Bioregion scale clustering

Across many bioregions time series clustering identified a ‘low’ cluster containing reefs
that suffer larger declines and generally have lower levels of coral cover than the other
clusters, a ‘medium’ cluster with intermediate levels of coral cover, and a ‘high’ cluster
with greater levels of cover recovery, with persistence to 2050 in some bioregions
(@sfig-access-esm1-5-bioregion-timeseries).

In many bioregions, median reef depth decreases as cluster coral cover decreases, with reefs
within the ‘high’ cluster commonly having the greatest median reef depths within the
bioregion (@fig-access-esm1-5-bioregion-depth).

![Median depth for reefs within each bioregion, separated into clusters assigned by
timeseries clustering. ‘Low’ clusters (green) are clusters with the lowest median
area-relative cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange)
are clusters with intermediate median area-relative cover between 2030 and 2060 within each
bioregion. ‘High’ clusters (blue) are clusters with the highest median area-relative cover
between 2030 and 2060 within each bioregion. Solid lines represent the median depth for each
cluster across reefs and bands represent the distribution of depth values within each
cluster. Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/depth_bioregion_violin.png){#fig-access-esm1-5-bioregion-depth
fig-align="center" width="14.82cm"}

In many bioregions reefs in the 'low' cover cluster have the lowest cover - DHW time series
correlation levels, while reefs in 'high' clusters have the highest correlation levels
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions in reef area-relative cover and mean time series DHW levels, with clusters having
similar mean DHW levels in many bioregions (@sfig-access-esm1-5-bioregion-dhwmean).
Similarly, there are no substantial differences between reef clusters in mean thermal
tolerance levels (@sfig-access-esm1-5-bioregion-dhwtol). There are no consistent differences
across bioregions and GCMs between clusters in their reefs' initial proportion of habitable
area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion). Across most
bioregions there are not consistent differences between clusters in connectivity metrics
such as source-to-sink connectivity ratio (@sfig-access-esm1-5-bioregion-logsosi) and total
connectivity strength (@sfig-access-esm1-5-bioregion-totalconnectivity).

### Partial Dependence

>>>>

In contrast to the broad temporal behaviour at the Management area scale, the relationship
between reef characteristics **\[TI: Ben, I think this is what you meant by "contextual
analysis factors"? depth, connectivity, dhw, carrying capacity etc?\]** and reefs' assigned
categories were found to be similar at the bioregion scale **\[TI: Ben, please confirm my
changes/edits are correct\]**. Results are therefore only shown and discussed for
ACCESS-ESM1-5, an intermediate Equilibrium Climate Sensitivity scenario [@sfig-ecs]. Those
interested in results for the other assessed GCMs are referred to Grier and Iwanaga
[-@grier2025]. Across bioregions, reefs in the High category not only had greater levels of
coral recovery but displayed persistence to 2050 [@sfig-access-esm1-5-bioregion-timeseries].
Reefs in the High category were consistently found to have the greatest median reef depths
across bioregions [@fig-access-esm1-5-bioregion-depth].

![Median depth for reefs within each bioregion, separated into clusters assigned by time
series clustering. ‘Low’ clusters (green) are clusters with the lowest median area-relative
cover between 2030 and 2060 within each bioregion. ‘Medium’ clusters (orange) are clusters
with intermediate median area-relative cover between 2030 and 2060 within each bioregion.
‘High’ clusters (blue) are clusters with the highest median area-relative cover between 2030
and 2060 within each bioregion. Solid lines represent the median depth for each cluster
across reefs and bands represent the distribution of depth values within each cluster.
Displays results for ACCESS-ESM1-5
GCM.](../figs/ACCESS-ESM1-5/bioregion/depth_bioregion_violin.png){#fig-access-esm1-5-bioregion-depth
fig-align="center" width="14.82cm"}

In many bioregions reefs in the Low category have the lowest correlation between coral cover
and DHW whereas reefs in the High category were found to have the strongest correlation
(@sfig-access-esm1-5-bioregion-dhwcor). Conversely, there are no consistent patterns across
bioregions when comparing their temporal behaviour with respect to their reef area, relative
cover, and mean of DHWs (@sfig-access-esm1-5-bioregion-dhwmean). There were no consistent
differences across bioregions and GCMs between clusters in their reefs' initial proportion
of habitable area occupied by corals (@sfig-access-esm1-5-bioregion-initialproportion).
Across most bioregions there were not consistent differences between clusters in
connectivity metrics such as source-to-sink connectivity ratio
(@sfig-access-esm1-5-bioregion-logsosi) and total connectivity strength
(@sfig-access-esm1-5-bioregion-totalconnectivity).

**\[TI: Note, I'm not sure if this section is interesting enough to keep now... let's
discuss.\]**
